Manuscript_Github_Pchabaudi_pir_TPMHeatmaps

### 2021-04-20 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir_TPMHeatmaps.Rmd
New chab heatmaps file

---
title: "Manuscript_Github_Pchabaudi_pir"
author: "Timothy Little"
date: "22/03/2021"
output: html_document
---

@@

## Figure 3c - Highest expressed all _pir_ heatmap

```{r highest_expressed_stage_heatmap_fig3b}

num_SorL <- cir_info %>% 
  group_by(SorL) %>% 
  summarise(num = n_distinct(Gene.ID))

labels_SorL <- setNames(paste0(num_SorL$SorL, ' (n = ', num_SorL$num, ')'),
                        num_SorL$SorL)

# Get order of pir genes by highest expressed genes in each sample.

combined_data_pir_ht_list <- lapply(names(twentyfour_avgtpm), 
       function(list_name){

         combined_data_pir <- getElement(twentyfour_avgtpm, 
                                     list_name)$tpm %>% 
           filter(Geneid %in% cir_id) %>% 
           column_to_rownames(var = 'Geneid')

         #Ordering that a more understandable order

         combined_data_pir <- combined_data_pir[,c(ordering(colnames(combined_data_pir)[-1], 
                                                            order_vector)),
                                                drop = FALSE]

         combined_data_ordered <- tibble(Geneid = names(apply(combined_data_pir, 1, mean)),
                                         means = apply(combined_data_pir, 1, mean),
                                         subfam = cir_info$SorL[match(names(apply(combined_data_pir, 1, mean)), 
                                                                            cir_info$Geneid)]) %>% 
           filter(means > 1) %>% 
           arrange(subfam, 
                   desc(means)) %>% 
           .$Geneid

         #Order the heatmap data by the order defined above (SorL, then mean expression).
         combined_data_pir_ht <- combined_data_pir[
           order(
             match(rownames(combined_data_pir),
                   combined_data_ordered),
             na.last = NA #Remove NAs
           ),
         ]

         #Change the colnames to the times
         colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
                                                                           pattern = '[[:digit:]]{2}h')

         #Order by time point
         combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
                                                      drop = FALSE] 
         #Add a column for the SorL
         combined_data_pir_ht$SorL <- cir_info$SorL[match(names(apply(combined_data_pir_ht, 1, mean)), 
                                                                cir_info$Geneid)]

         return(combined_data_pir_ht)
       })
names(combined_data_pir_ht_list) <- names(twentyfour_avgtpm)

maxcol <- ceiling(
  max(
    unlist(
      lapply(combined_data_pir_ht_list,
             function(list_element){
               max(
                 log2(1+list_element[!colnames(list_element) %in% 'SorL'])
               )
             }
      )
    )
  )
)

lapply(names(combined_data_pir_ht_list),
       function(list_name_to_order){

         combined_data_pir_ht <- getElement(combined_data_pir_ht_list,
                                            list_name_to_order)

         pir_highest_tpm_order <- rownames(combined_data_pir_ht)

         #Sub-family annotation
         clade_info <- combined_data_pir_ht$SorL
         annotation_col <- SorL_col

         #Subfamily annotation

         clade_ha <- rowAnnotation(
           'sub-family' = anno_simple(
             clade_info,
             col = annotation_col,
           ),
           annotation_name_gp = gpar(fontsize = 10)
         )

         anc_ha <- rowAnnotation(foo = anno_mark(at =  na.omit(match(clade_info, 'ancestral')) , 
                                                 labels = 'ancestral',
                                                 labels_gp = gpar(fontsize = 7),
                                                 link_width = unit(3, "mm")))

         #lapply across the datasets now, for each order
         ht_list <- lapply(names(combined_data_pir_ht_list),
                           function(list_name){

                             combined_data_pir_ht <- getElement(twentyfour_avgtpm, 
                                                                list_name)$tpm %>% 
                               column_to_rownames(var = 'Geneid')

                             #Get the times instead of the stage names.
                             colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
                                                                           pattern = '[[:digit:]]{2}h')

                             combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
                                                                          drop = FALSE]

                             Heatmap(
                               log2(1+combined_data_pir_ht)[pir_highest_tpm_order,],
                               cluster_rows = FALSE,
                               cluster_columns = FALSE,
                               show_row_names = FALSE,
                               name = 'log(TPM+1)', 
                               col = colorRamp2(c(0,
                                                  maxcol/4,
                                                  maxcol/2,
                                                  3*maxcol/4,
                                                  maxcol),
                                                viridis(5)),
                               na_col = 'white', 
                               border = TRUE,
                               border_gp = gpar(lwd = 0.1),
                               column_names_gp = gpar(fontsize = 10),
                               row_dend_width = unit(3, "cm"),
                               width = unit(4.5, 'cm'),
                               height = unit(7,'cm'),
                               row_title = NULL,
                               left_annotation = if(
                                 list_name %in% dplyr::first(names(combined_data_pir_ht_list))
                               ){
                                 clade_ha
                               }else{
                                 NULL
                               },
                               right_annotation = if(
                                 list_name %in% dplyr::last(names(combined_data_pir_ht_list))
                               ){
                                 anc_ha
                               }else{
                                 NULL
                               },
                               # row_split =  cir_info$SorL[match(pir_highest_tpm_order, 
                               #                                        cir_info$Geneid)],
                               column_title = paste(list_name),
                               heatmap_legend_param = list(
                                 labels_gp = gpar(fontsize = 10),
                                 title_gp = gpar(fontsize = 10,
                                                 fontface = 'bold'),
                                 at = c(0,
                                        maxcol/4,
                                        maxcol/2,
                                        3*maxcol/4,
                                        maxcol)
                               )
                             )
                           }
         )
         ht_list <- Reduce(`+`, ht_list)
         clade_lgd <- Legend(
           title = 'sub-family',
           at = labels_SorL,
           legend_gp = gpar(fill = SorL_col)
         )

         if(global_save_var){        
           pdf('Manuscript_figure3c_heatmap.pdf'
               , width = 10
               , height = 4
           )
           draw(ht_list, 
                row_title_side = 'left',
                column_title_side = 'bottom',
                heatmap_legend_side = 'right',
                annotation_legend_side = 'right',
                annotation_legend_list = list(clade_lgd),
                ht_gap = unit(.35,'cm'),
                merge_legend = TRUE)
           dev.off()
         }

         return(
           draw(ht_list, 
                row_title_side = 'left',
                column_title_side = 'bottom',
                heatmap_legend_side = 'right',
                annotation_legend_side = 'right',
                annotation_legend_list = list(clade_lgd),
                ht_gap = unit(.35,'cm'),
                merge_legend = TRUE
           ))
       })

## 2D plot of PC1 v PC2 - Supp figure 4A

```{r pca_calculation, results = TRUE}
twentyfour_tpm <- twentyfour_rMT
rownames_geneid <- twentyfour_tpm$Geneid
twentyfour_tpm_forpca <- select(twentyfour_tpm, -Geneid)
rownames(twentyfour_tpm_forpca) <- rownames_geneid
twentyfour_tpm_pca <- prcomp(twentyfour_tpm_forpca, center = TRUE, scale. = TRUE)
#Need to look into what exactly scale and center are doing
```

```{r 2d_pca, results = TRUE}
twentyfour_tpm_pca_var <- get_pca_var(twentyfour_tpm_pca)
twentyfour_tpm_pca_coords <- data.frame(twentyfour_tpm_pca_var$coord)

twentyfour_tpm_pca_coords <- data.frame(sample = rownames(twentyfour_tpm_pca_var$coord),
                                        PC1 = twentyfour_tpm_pca_coords$Dim.1,
                                        PC3 = twentyfour_tpm_pca_coords$Dim.3,
                                        PC2 = twentyfour_tpm_pca_coords$Dim.2) %>% 
  dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
  dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,1])

colornames <- viridis(length(unique(twentyfour_tpm_pca_coords$time)))
colornames <- setNames(colornames, 
                       sort(unique(twentyfour_tpm_pca_coords$time)))

pca_importance <- as.data.frame(summary(twentyfour_tpm_pca)$importance)
pc1_importance <- round(pca_importance$PC1[2]*100, digits = 1)
pc2_importance <- round(pca_importance$PC2[2]*100, digits = 1)
pc3_importance <- round(pca_importance$PC3[2]*100, digits = 1)

(
  plot <- ggplot(twentyfour_tpm_pca_coords,
                 aes(
                   x = PC1,
                   y = PC2,
                   col = time
                 )) +
    geom_point(size = 3,
               alpha = 0.8) +
    scale_color_manual(values = colornames) +
    xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
    ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
    guides(shape=guide_legend(ncol=2),
           col = guide_legend(ncol=2)) +
    theme_classic() +
    theme(legend.key.size = unit(4, 'mm'),
          legend.text = element_text(size = 7))
)

if(global_save_var){
  ggsave2(filename = 'Manuscript_SuppFigure3_chabaudi24h_pca_nopseudo.png',
          plot = plot,
          dpi = 300,
          width = 15,
          height = 10,
          units = 'cm')
}

# Calculate and cluster with z-scores {.tabset}

z-score represents the number of standard deviations from the mean of the data. To calculate from TPM or counts for a given gene X:

1. Convert the count/RPKM values of each gene into log values.
2. Calculate the mean and standard deviation of X gene log values in each sample
3. For first sample: (gene X log value - mean of log values of samples)/ standard deviation of log values of all samples
4. Now. i have the z-score for gene x in first sample. Using the above protocol, I can convert all genes log values into z-score.

* Need to make sure that the z-score is calculated within each transmission method!

```{r zscores}
tpm_to_zscore = function(input){

  #If the input does not have a TPM element then assume that the input itself is the TPM object
  if(any(names(input) %in% 'tpm')){
    tpm_object <- input$tpm
  }else{
    tpm_object <- input
    warning('No \'tpm\' element in the supplied input so function will assume that the input is the tpm data')
  }

  #Want all of them so will just ignore Geneid
  input_logged <- tpm_object %>%
    mutate_at(vars(-matches('Geneid')), 
              log1p)

  gene_mean_scores = input_logged %>% 
    melt(id.vars = 'Geneid') %>% 
    group_by(Geneid) %>% 
    dplyr::summarise(mean = mean(value),
                     std_dev = sd(value))

  gene_vector = as.character(input_logged$Geneid)

  input_z = lapply(gene_vector, function(gene){
    mean_gene = filter(gene_mean_scores, 
                       Geneid == paste(gene)) %>% 
      select(mean)
    std_gene = filter(gene_mean_scores, 
                      Geneid == paste(gene)) %>% 
      select(std_dev)
    gene_scores = input_logged %>% 
      filter(Geneid == paste(gene)) %>% 
      select(-Geneid)
    if(std_gene != 0){
      gene_z = sapply(gene_scores, function(score){(score-mean_gene)/std_gene}) %>% data.frame
    }else{
      gene_z = (sapply(gene_scores, function(score){0})) %>% data.frame %>% t
    }
    colnames(gene_z) = colnames(gene_scores) #For some reason apply makes the column names weird
    gene_z = cbind(data.frame(Geneid = paste(gene)), gene_z)
  })

  #Use reduce and rbind to turn the list into one df
  input_z_df = purrr::reduce(input_z, 
                             rbind)
  return(list(zscores = input_z_df,
              tpm = tpm_object,
              experiment = input$experiment))
}

#Make and save z-scores
twentyfour_z <- lapply(twentyfour_avgtpm,
                       tpm_to_zscore)

###
```

# Phaseograms of all the genes.

```{r row_order_phaseograms}

#Getting the row order for each phaseogram for each experiment
# Want the data ordered by the highest zscore for each gene in each stage.
phaseogram_row_order <- lapply(twentyfour_z, 
                               function(list_elements){
                                 zscore_df <- list_elements$zscores
                                 zscore_ordered <- melt(zscore_df, 
                                                        id = 'Geneid',
                                                        variable.name = 'stage',
                                                        value.name = 'zscore') %>% 
                                   group_by(Geneid) %>% 
                                   top_n(zscore, n = 1) %>% 
                                   arrange(stage, 
                                           desc(zscore)) %>%
                                   distinct(Geneid, 
                                            .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
                                   filter(zscore > 0) #For some reason some remain with zero zscore
                                 # zscore_ordered <- zscore_ordered[order(
                                 #   match(
                                 #     zscore_ordered$stage, 
                                 #     order_vector
                                 #   )
                                 # ),]
                                 return(c(list_elements, 
                                          list(zscore_order = zscore_ordered)))
                               })

phaseogram_time_order <- lapply(names(phaseogram_row_order), 
                                function(list_element){
                                  transmission_list <- getElement(phaseogram_row_order,
                                                                  list_element)
                                  order_by_time <- transmission_list$zscore_order %>% 
                                    mutate(stage = times_stages_table$time[match(stage,
                                                                                 times_stages_table$r_stage_diff_count)]) %>% 
                                    arrange(stage)
                                  return(list(zscore_order = order_by_time))
                                }
)
names(phaseogram_time_order) <- names(phaseogram_row_order)

```{r function_draw_heatmap_zscore}

draw_heatmap_zscore <- function(zscore_list,
                                title = '',
                                gene_ordering_vector = NULL,
                                filter_genes_vector = NULL,
                                extra_title = NULL,
                                annotation_clade = NULL,
                                annotation_col,
                                threshold_tpm = NULL,
                                use_times = FALSE,
                                use_order_vector = TRUE,
                                tpm_mat = TRUE,
                                tpm_breaks = NULL,
                                ...){
  if(is.null(tpm_breaks) & tpm_mat){
    #Need the generate the breaks in the data for the heatmap legend if not provided.
    max_tpm <- lapply(zscore_list, 
                      function(list_element){
                        #list_element <- zscore_list[[1]]
                        heatmap_data <- as.data.frame(list_element$zscores)
                        title <- list_element$experiment

                        if(!'Geneid' %in% colnames(heatmap_data)){
                          stop('No Geneid column in the heatmap data.frame')
                        }

                        if(!is.null(filter_genes_vector)){
                          heatmap_data <- as.data.frame(
                            filter(heatmap_data, 
                                   Geneid %in% filter_genes_vector)
                          )
                        }

                        rownames(heatmap_data) <- as.character(heatmap_data$Geneid)

                        heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
                        #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!

                        if(!is.null(gene_ordering_vector)){
                          #Remove genes not in the main data frame but in the ordering vector, 
                          # and then order based on the vector
                          remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
                          heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
                          heatmap_data <- heatmap_data[na.omit(
                            match(gene_ordering_vector, 
                                  rownames(heatmap_data))
                          ),,
                          drop=FALSE]
                        }

                        geneids <- rownames(heatmap_data)

                        #Get max/min TPM scores for the side annotation heatmap                       
                        tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
                                                          list_element$tpm$Geneid),] %>% 
                          select(-Geneid) %>% 
                          as.matrix %>% 
                          log1p
                        rownames(tpm_mat) <- geneids

                        tpm_min <- rowMins(tpm_mat)
                        names(tpm_min) <- geneids
                        tpm_max <- rowMaxs(tpm_mat)
                        names(tpm_max) <- geneids
                        return(list(tpm_max = tpm_max,
                                    tpm_min = tpm_min))
                      })

    tpm_max_all <- mround(
      max(
        unlist(
          lapply(max_tpm, 
                 getElement,
                 'tpm_max')
        )
      ), 
      base = 1)

    tpm_breaks <- round(
      c(0,
        tpm_max_all/4,
        tpm_max_all/2,
        tpm_max_all*(3/4),
        tpm_max_all),
      digits = 2
    )
  }

  #Generate the colour function from the TPM legend breaks for the TPM annotation
  if(tpm_mat){
    tpm_col_fun <- colorRamp2(
      breaks = tpm_breaks,
      colors = viridis(length(tpm_breaks))
    )
  }

  if(is.numeric(threshold_tpm)){
    #Apply a threshold to the TPM in order to remove low expressed genes.
    # Remember to add log1p here! 
    # Threshold uses the max value, so it ony needs to be above thresh 
    #   in one stage avg in one transmission method.
    #Will alter the gene_filter vector in order to only select the genes above threshold

    gene_maxs <- bind_rows(
      lapply(max_tpm, 
             getElement,
             'tpm_max')
    )

    remove_genes <- names(gene_maxs)[apply(
      gene_maxs, 
      2, 
      function(maxs){
        all(maxs < log2(1+threshold_tpm))
      }
    )]

    filter_genes_vector <- filter_genes_vector[!filter_genes_vector %in% remove_genes]

  if(is.character(annotation_clade)){
    order_cir <- gene_ordering_vector[gene_ordering_vector %in% filter_genes_vector]
    clade_info <- getElement(cir_info,
                             annotation_clade)[match(order_cir, 
                                                     cir_info$Geneid)]
    names(clade_info) <- str_split_fixed(order_cir, 
                                         pattern = '_', 
                                         n = 2)[,2]
    annotation_col <- annotation_col
  }

  heatmaps <- lapply(zscore_list, 
                     function(list_element){
                       #list_element <- zscore_list[[1]]
                       heatmap_data <- as.data.frame(list_element$zscores)
                       title <- list_element$experiment

                       if(!'Geneid' %in% colnames(heatmap_data)){
                         stop('No Geneid column in the heatmap data.frame')
                       }

                       if(!is.null(filter_genes_vector)){
                         heatmap_data <- as.data.frame(
                           filter(heatmap_data, 
                                  Geneid %in% filter_genes_vector)
                         )
                       }

                       rownames(heatmap_data) <- as.character(heatmap_data$Geneid)

                       heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
                       #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!

                       if(!is.null(gene_ordering_vector)){
                         #Remove genes not in the main data frame but in the ordering vector,
                         #  and then order based on the vector
                         remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
                         heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
                         heatmap_data <- heatmap_data[na.omit(
                           match(gene_ordering_vector, 
                                 rownames(heatmap_data))
                         ),,
                         drop=FALSE]
                       }

                       if(tpm_mat){
                         #Get max/min TPM scores for the side annotation heatmap                       
                         tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
                                                           list_element$tpm$Geneid),] %>%
                           select(-Geneid) %>% 
                           as.matrix %>% 
                           log1p

                         tpm_min <- rowMins(tpm_mat)
                         tpm_max <- rowMaxs(tpm_mat)

                         names(tpm_min) = rownames(heatmap_data)
                         names(tpm_max) = rownames(heatmap_data)

                         tpm_ha <- rowAnnotation(
                           min.max = anno_simple(cbind(tpm_min, 
                                                       tpm_max), 
                                                 col = tpm_col_fun,
                                                 gp = gpar(col = 'black',
                                                           lwd = 0.1),
                                                 border = TRUE,
                                                 simple_anno_size = unit(0.2, "cm")),
                           show_annotation_name = FALSE)
                       }

                       #Draw the left annotation is needed, only for the first element.

                       if(
                         identical(list_element,
                                   dplyr::first(zscore_list)) & is.character(annotation_clade)
                       ){
                         clade_ha <- rowAnnotation(
                           'subfamily' = anno_simple(
                             clade_info,
                             col = annotation_col
                           )
                         )
                       }

                       #Change rownames to remove PCHAS_
                       rownames(heatmap_data) = str_split_fixed(rownames(heatmap_data), 
                                                                pattern = '_', 
                                                                n = 2)[,2]

                       if(use_times){
                         colnames(heatmap_data) <- str_extract(colnames(heatmap_data), 
                                                               pattern = '[[:digit:]]{2}h')
                       }

                       if(use_order_vector){
                         #Order the columns by ordering_vector
                         heatmap_data <- heatmap_data[,order(match(colnames(heatmap_data), 
                                                                   order_vector)),
                                                      drop = FALSE]
                       }else{
                         heatmap_data <- heatmap_data[,order(colnames(heatmap_data)),
                                                      drop = FALSE] 
                       }

                       h = Heatmap(heatmap_data,
                                   name = paste0('z-score',
                                                 extra_title),
                                   col = c(cbPalette[6],
                                           'white', 
                                           cbPalette[7]),
                                   row_names_gp = gpar(fontsize = ifelse(
                                     nrow(heatmap_data) > 13, 
                                     12*(13/nrow(heatmap_data)), 
                                     12)
                                   ),
                                   column_names_gp = gpar(fontsize = 10),
                                   row_dend_width = unit(3, "cm"),
                                   right_annotation = if(tpm_mat){
                                     tpm_ha
                                   }else{
                                     NULL
                                   },
                                   left_annotation = 
                                     if(identical(list_element,
                                                  dplyr::first(zscore_list)) & is.character(annotation_clade)){
                                       clade_ha
                                     }else{
                                       NULL
                                     }, 
                                   #Only want the clade annotation when it is first heatmap
                                   width = unit(4.5, 'cm'),
                                   cluster_columns = FALSE,
                                   height = unit(7,'cm'),
                                   row_title = NULL,
                                   cluster_rows = FALSE,
                                   column_title = paste(title),
                                   heatmap_legend_param = list(
                                     at = c(3,2,1,0,-1,-2,-3),
                                     labels_gp = gpar(fontsize = 10),
                                     title_gp = gpar(fontsize = 10)
                                   ),
                                   ...
                       )
                       return(h)
                     })

  ht_list <- Reduce('+', heatmaps) #To add the list elements together

  if(tpm_mat){
    #Draw the legend
    tpm_lgd <- Legend(
      title = "Min/Max log(TPM + 1)", 
      col = tpm_col_fun, 
      at = tpm_breaks, 
      labels = tpm_breaks,
      title_gp = gpar(fontsize = 7)
    )
  }

  if(is.character(annotation_clade)){
    clade_lgd <- Legend(
      title = 'sub-family',
      at = names(annotation_col),
      legend_gp = gpar(fill = annotation_col)
    )
  }

  draw(ht_list, 
       row_title_side = 'left',
       column_title_side = 'bottom',
       heatmap_legend_side = 'left',
       #Only include the annotation legends for the annotations we actually have.
       annotation_legend_list = if(is.character(annotation_clade) & tpm_mat){
         list(tpm_lgd,
              clade_lgd)
       }else if(tpm_mat & !is.character(annotation_clade) ){
         list(tpm_lgd)
       } else if(!tpm_mat & is.character(annotation_clade) ){
         list(clade_lgd)
       })
}
```

## Supplementary figure 5 - Phaseograms of the entire genome.

```{r phaseograms_whole_genome}
if(global_save_var){
  pdf(file = 'Manuscript_SuppFig5_phaseogram_allgenes.pdf',
      height = 5,
      width = 5)
  draw_heatmap_zscore(zscore_list = twentyfour_z,
                      gene_ordering_vector = phaseogram_time_order$RMT$zscore_order$Geneid,
                      use_times = TRUE,
                      use_order_vector = FALSE,
                      show_row_names = FALSE,
                      tpm_mat = FALSE)
  dev.off()
}

draw_heatmap_zscore(zscore_list = twentyfour_z,
                    gene_ordering_vector = phaseogram_time_order$RMT$zscore_order$Geneid,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_mat = FALSE)

##  Supplementary figure 6 - Phaseograms of the _cir_ genes with clade annotations

```{r pir_gene_phaseogram_breaks}
  max_tpm <- lapply(twentyfour_z, 
                    function(list_element){
                      #list_element <- zscore_list[[1]]
                      heatmap_data <- as.data.frame(list_element$zscores)
                      title <- list_element$experiment

                      if(!'Geneid' %in% colnames(heatmap_data)){
                        stop('No Geneid column in the heatmap data.frame')
                      }

                            heatmap_data <- as.data.frame(
                              filter(heatmap_data, 
                                     Geneid %in% cir_id)
                            )

                          rownames(heatmap_data) <- as.character(heatmap_data$Geneid)

                          heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
                          #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
                          # 
                          # if(!is.null(gene_ordering_vector)){
                          #   #Remove genes not in the main data frame but in the ordering vector, 
                          #   # and then order based on the vector
                          #   remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
                          #   heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
                          #   heatmap_data <- heatmap_data[na.omit(
                          #     match(gene_ordering_vector, 
                          #           rownames(heatmap_data))
                          #   ),,
                          #   drop=FALSE]
                          # }

                          geneids <- rownames(heatmap_data)

                          #Get max/min TPM scores for the side annotation heatmap                       
                          tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
                                                            list_element$tpm$Geneid),] %>% 
                            select(-Geneid) %>% 
                            as.matrix %>% 
                            log1p
                          rownames(tpm_mat) <- geneids

                          tpm_min <- rowMins(tpm_mat)
                          names(tpm_min) <- geneids
                          tpm_max <- rowMaxs(tpm_mat)
                          names(tpm_max) <- geneids
                          return(list(tpm_max = tpm_max,
                                      tpm_min = tpm_min))
                        })

  tpm_max_all <- mround(
    max(
      unlist(
        lapply(max_tpm, 
               getElement,
               'tpm_max')
      )
    ), 
    base = 1)

  tpm_breaks <- round(
    c(0,
      tpm_max_all/4,
      tpm_max_all/2,
      tpm_max_all*(3/4),
      tpm_max_all),
    digits = 2
  )
```

```{r phaseograms_cir}

if(global_save_var){
 pdf(file = 'Manuscript_SuppFig6_phaseogram_cirgenes.pdf',
      height = 5,
      width = 5)
  draw_heatmap_zscore(zscore_list = twentyfour_z,
                      gene_ordering_vector = phaseogram_time_order$RMT$zscore_order$Geneid,
                      filter_genes_vector = cir_id,
                      annotation_clade = 'SorL',
                      annotation_col = SorL_col,
                      threshold_tpm = 1,
                      use_times = TRUE,
                      use_order_vector = FALSE,
                      show_row_names = FALSE,
                      tpm_mat = FALSE)
  dev.off()
}

draw_heatmap_zscore(zscore_list = twentyfour_z,
                    gene_ordering_vector = phaseogram_time_order$RMT$zscore_order$Geneid,
                    filter_genes_vector = cir_id,
                    annotation_clade = 'SorL',
                    annotation_col = SorL_col,
                    threshold_tpm = 1,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_mat = FALSE)

@@

# Supp figure 9 - Pir expression of AAPLs and ChAPLs.

```{r Supp_fig_AapL-ChaAPl_expression}

cir_plot <- twentyfour_rMT %>% 
  filter(Geneid %in% cir_id) %>%
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(
    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,2],
    time = str_split_fixed(sample, pattern = '_', n = 3)[,1],
    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)],
    chapl_aapl = factor(cir_info$chapl_aapl_name[match(Geneid, 
                                                       cir_info$Geneid)],
                        levels = sort(unique(cir_info$chapl_aapl_name[match(Geneid, 
                                                                            cir_info$Geneid)])))) %>% 
  filter(!(time %in% c('sch', 
                       'Unknown')),
         !is.na(chapl_aapl) ) %>% 
  group_by(Geneid) %>% 
    arrange(desc(tpm))

  cir_plot <- mutate(cir_plot, 
                     Geneid = factor(Geneid, 
                                     levels = unique(cir_plot_order$Geneid)))
  cir_plot_chapl_aapl_supp_info <- cir_plot

  max_aapl_chapl <- ceiling(max(log2(cir_plot$tpm+1)))

  breaks = seq(0, max_aapl_chapl, by = 1)
    limits = c(0, max_aapl_chapl)
    point_size = 0.25

  lapply(
    sort(unique(cir_plot$time)),
    function(time_vec){

    cir_plot <- filter(cir_plot, 
                       time == time_vec)

      cir_plot$tpm <- log2(1+cir_plot$tpm)

    cir_plot$Geneid <- factor(
      str_remove(cir_plot$Geneid, 
                 pattern = 'PCHAS_'),
      levels = str_remove(levels(cir_plot$Geneid), 
                          pattern = 'PCHAS_')
    )

    y <- ggplot(
      data = cir_plot,
      mapping = aes(
        x = Geneid,
        y = tpm
      )
    ) +
      stat_summary(fun.y = median, 
                   fun.ymin = median, 
                   fun.ymax = median,
                   geom = "bar", 
                   width = 0.75, 
                   size = 0.25, 
                   position = 'dodge',
                   fill = 'lightblue') +
      geom_point(stat = 'identity', 
                 position = position_dodge(width = 0.75),
                 size = point_size) +
      scale_y_continuous(breaks = breaks,
                         limits = limits) +
      scale_fill_manual(values = c('skyblue','red')) +
      ylab('log(TPM +1)') +
      facet_grid(.~chapl_aapl, 
                 scales = "free_x",
                 space = "free_x") +
      theme_classic() +
      theme(
        axis.text.x = element_text(angle = 90,
                                   size = 5)
      ) +
      ggtitle(paste(unique(cir_plot$time)))

    if(global_save_var){
      ggsave2(
        paste0('Pchabaudi_sibe-by-side_chapl_aapl_',
               time_vec,
               '.png'),
        plot = y,
        dpi = 300,
        width = 25,
        height = 10,
        units = 'cm'
      )
    }

    return(y)
    })
```

>>>>

---
title: "Manuscript_Github_Pchabaudi_pir_TPMHeatmaps"
author: "Timothy Little"
date: "20/04/2021"
output: html_document
---

@@

```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@
